From Education To Entrepreneurship

### 2014-11-29 FromEducationToEntrepreneurship.Rmd
Research Interest, Question and Literature Review

>>>>

*Introduction to Collaborative Social Science Data Analysis*

### Research Interest

The process of allocating public funds for education is an opportunity to make calculated long-term investments in a nation's populace.  The rationale for decisions made during this process, however, is often difficult to understand.  For example, while many claims have been made in entrepreneurship literature regarding the importance of cultural and educational factors -- not just structural and economic factors -- in determining individual and societal tendencies to start a business, these factors are often not included in the decisionmaking process. As a matter of fact, only a few empirical studies have actually been carried out on this topic and most of them focused on which attitudes were favorable to starting a business, while almost none explored how cultural patterns affected these dispositions.

When cultural factors are examined, they are considered as a given feature that can hardly be traced. By having a closer look at education systems, however, we intend to explore the processes through which some of these cultural patterns are enhanced and passed from one generation to the next. Aside from parental education, schools are indeed the very place where children are taught how to fit into society. This research is an attempt to analyze  one trait in particular, self-confidence.

### Research Question

Drawing a comparison between France and Britain, to what extent do the behavioral characteristics of the education system - individualistic, group-oriented, and more or less positive reinforcement in teaching - affect the rate of entrepreneurship in each country?

### Literature Review

What determines entrepreneurship at a country-level has been long searched already but although economic, demographic or geographic factors have been looked at extensively, much less has been done regarding cultural or educational factors. However, as Freytag and Thurik stated in 2007, “the relative stability of differences in entrepreneurship across countries suggests that factors other than economic ones are at play” [@Freytag2007] . Even if “in recent years research has increasingly devoted itself to the subject” [@OECD2009], a lot still needs to be done, as the few studies that explored the relationship between attitudes and foundation activities barely searched what determined these attitudes.

Over a hundred years has passed since Max Weber tried to explain how religious-ethical motivations affect entrepreneurship [@weberprotestant], and religion may be less effective in justifying why some are more likely to start businesses than others. Still, that was a founding stone and it is now widely admitted that “cultural features influence attitudes towards start-ups and that these attitudes, in turn, have an impact on start-up activities” [@OECD2009].

Despite various claims of the existence of a cultural effect, few empirical studies have actually been carried out. It took a few decades more before Davidsson, along with Delmar at first (1992), then alone (1995), and then with Wiklund [@davidsson1997values], at his turn, laid a stepping stone by investigating the relationship between structural factors -- mainly economic, and cultural factors. As they compared two Swedish regions similar in terms of structure and observed different outcomes, they concluded there may be a small cultural effect in attitudes toward start-up activities.

This thesis is an attempt at finding out whether different patterns in education lead to different outcomes in terms of start-up activities. Indeed, like many in the education literature tend to assert, we believe school personnel have an affect on student self-esteem [@Scott1996]. We will further explore how this mechanism works. As for entrepreneurship literature, it generally acknowledges the influence of self-confidence on start-up activities, although “self-confidence in one’s own capabilities to start a business has been found to depend almost exclusively on the individual features of the respective person and his or her integration into social networks” [@OECD2009]. We intend to show that these individual characteristics are partly shaped at an institutional and collective level through the education systems and staff attitudes.

Of course, King and Sobel [@Kerry2008], among others, have already tried to show how school choice can increase the rate of youth entrepreneurship, but their perspective was slightly different. They  asserted that a more business-like environment in K-12 education was the key, which was favored by voucher programs. We believe there is a more widely spread systematic effect of cultural and educational patterns. David Watkins [@Watkins2010] cross-cultural perspective on teaching and learning shall be useful in that respect. Kai-Ming Cheng and Kam-Cheung Wong’s study about school effectiveness in East Asia [@Cheng2006] could also be of help.

Also, as self-confidence is considered to rely partly on one’s integration within social networks [@OECD2009], we will refer to cross-cultural studies such as that of Green, Deschamps and Páez [@green2005variation] about individualism and collectivism, what Freeman [@freeman1997demographic] called idiocentrism and allocentrism. Raeff, Greenfield and Quiroz [-@raeff2000conceptualizing] have also conceptualized interpersonal relationships in both contexts. More school-focused, Yamauchi [@yamauchi1998individualism] as well as Davies and Aurini [@davies2003homeschooling] have looked at individualism and collectivism in the classroom and in pedagogy respectively. All these will be important on the way.

Besides direct or indirect school effects, we will necessarily look into entrepreneurship and its determinants at large. Thus we will quote Grilo and Thurik’s research [@Grilo2005b] about what determines entrepreneurship engagement levels throughout the European Union, and also Wennekers Sander, Ulhaner Lorraine and Thurik Roy’s explanation of variation across countries [@Wennekers2002]. Eventually, with Gartner [@Gartner1990] we will have a closer look at entrepreneurship’s underlying meanings in research and, along with Acs, Desai and Klapper [@Acs2008], we will ask ourselves what “entrepreneurship” data really shows.
